Debiased Nonparametric Regression Estimation

### 2024-05-13 src/explore/20240506-npregest.qmd
Fixed multisource simulation 2. It works, but I need to change the estimator to be a GLS estimate instead of a simple average.

library(dplyr)
library(doParallel)
library(doRNG)
library(npreg)
library(mgcv)
library(randomForest)

@@

gen_pop <- function(N_obs, n1_obs) {
  x <- runif(N_obs, 0, 8 * pi)
  eps <- rnorm(N_obs)
  y <- 0.2 * x + sin(x) + 0.3 * eps

  pi1 <- n1_obs / N_obs
  z <- (x - mean(x)) / sd(x)
  pi2 <- ifelse(pnorm(z - 0.5) > 0.7, 0.7, pnorm(z - 0.5))
  pi2 <- ifelse(pi2 < 0.02, 0.02, pi2)

  return(tibble(X = x, Y = y, pi1 = pi1, pi2 = pi2))

@@

  # 1. Get mhat.
  mx <- 0.2 * p1_df$X + sin(p1_df$X)
  mx2 <- 0.2 * p2_df$X + sin(p2_df$X)

@@

  # 1. Estimate mhat.
  if (type == "loess") {
    mod <- loess(Y ~ X, p2_df, control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    mod <- randomForest(Y ~ X, p2_df)
  } else if (type == "spline") {
    # mod <- ss(x = p2_df$X, y = p2_df$Y)
    # mx <- predict(mod, x = p1_df$X)$y
    # mx2 <- predict(mod, x = p2_df$X)$y
    mod <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2_df)
  } else if (type == "linear") {
    mod <- lm(Y ~ X, data = p2_df)
  }

@@

  p1_df <- p1_df %>%
    left_join(p2_df, by = join_by(X, Y, pi1, pi2, del1, del2)) %>%
    mutate(del2a = ifelse(is.na(del2a), 0, del2a)) %>%
    mutate(del2b = ifelse(is.na(del2b), 0, del2b))

  # 1. Estimate mhat in both splits
  if (type == "loess") {
    moda <- loess(Y ~ X, p2a_df, control = loess.control(surface="direct"))
    modb <- loess(Y ~ X, p2b_df, control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    moda <- randomForest(Y ~ X, p2a_df)
    modb <- randomForest(Y ~ X, p2b_df)
  } else if (type == "spline") {
    # moda <- ss(x = p2a_df$X, y = p2a_df$Y)
    # modb <- ss(x = p2b_df$X, y = p2b_df$Y)
    # ma_2 <- predict(moda, x = p2_df$X)$y
    # mb_2 <- predict(modb, x = p2_df$X)$y
    # ma_1 <- predict(moda, x = p1_df$X)$y
    # mb_1 <- predict(modb, x = p1_df$X)$y
    # mx <- (ma_1 + mb_1) / 2
    moda <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2a_df)
    modb <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2b_df)
  } else if (type == "linear") {
    moda <- lm(Y ~ X, data = p2a_df)
    modb <- lm(Y ~ X, data = p2b_df)
  }

@@

  # p2a_df <- p2a_df %>%
  #   mutate(pi2a = pi2 * nrow(p2a_df) / nrow(p2_df))
  # p2b_df <- p2b_df %>%
  #   mutate(pi2b = pi2 * nrow(p2b_df) / nrow(p2_df))

  # v1_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
  # v2_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))
  # cov_t <- matrix(NA, nrow = nrow(p2_df), ncol = nrow(p2_df))

  # # Since Phase 1 is SRS
  # p1ij <- nrow(p1_df) / N_obs * (nrow(p1_df) - 1) / (N_obs - 1)
  # eps_a <- p2_df$Y - ma_2
  # eps_b <- p2_df$Y - mb_2

  # n2a <- nrow(p2a_df)
  # n2b <- nrow(p2b_df)
  # n2 <- nrow(p2_df)

  # for (i in 1:nrow(p2_df)) {
  #   for (j in 1:nrow(p2_df)) {
  #     if (i == j) {
  #       v1_t[i, j] <- 
  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_b[i]^2 / p2_df$pi2[i]^2
  #       v2_t[i, j] <- 
  #       1 / 4 * (2 - p2_df$pi2[i]) / p2_df$pi1[i] * eps_a[i]^2 / p2_df$pi2[i]^2
  #       cov_t[i, j] <- 
  #       1/(p2_df$pi2[i] * p2_df$pi1[i]) * n2a/n2 * (n2a - 1)/n2 * eps_b[i] * eps_a[i]
  #     } else {
  #       v1_t[i, j] <- 1 / 4 * (p2_df$pi2[i]) / (p2_df$pi1[i] * p2_df$pi1[j]) *
  #         eps_b[i] * eps_b[j] / (p2_df$pi2[i] * p2_df$pi2[j])
  #       v2_t[i, j] <- 1 / 4 * 1 / (p2_df$pi1[i] * p2_df$pi1[j]) *
  #         eps_a[i] * eps_a[j] / (p2_df$pi2[i] * p2_df$pi2[j])
  #       cov_t[i, j] <- 1 / (p2_df$pi2[i] * p2_df$pi1[i]) * n2a / n2 * 
  #       ((n2 - n2a) / (n2 - 1) + (n2a - 1)/n2) * eps_b[i] * eps_a[i]
  #     }
  #   }
  # }

  # v1 <- sum(as.numeric(v1_t)) / N_obs^2
  # v2 <- sum(as.numeric(v2_t)) / N_obs^2
  # v3 <- sum(as.numeric(cov_t)) / N_obs^2

  # vhat <- (v0 + v1 + v2 + v3)

@@

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "randomForest", "mgcv")) %dopar% {

@@

>>>>

library(dplyr)
library(doParallel)
library(doRNG)
library(mgcv)
library(randomForest)

@@

gen_pop <- function(N_obs, n1_obs) {
  x <- runif(N_obs, 1, 2)
  z <- rnorm(N_obs)
  eps <- rnorm(N_obs)
  # y <- 0.05 * x^2 + 2 * eps + exp(x - max(x) + 4) + 2 * sin(x)
  y <- x + z + eps

  pi1 <- n1_obs / N_obs
  pi2 <- ifelse(pnorm(z - 1) > 0.7, 0.7, pnorm(z - 1))
  pi2 <- ifelse(pi2 < 0.02, 0.02, pi2)

  w1 <- 1 / pi1
  w2 <- 1 / pi2

  return(tibble(X = x, Y = y, pi1 = pi1, pi2 = pi2, w1 = w1, w2 = w2))

@@

  # 1. Get mhat.
  mx <- p1_df$X
  mx2 <- p2_df$X

@@

  # 1. Estimate mhat.
  if (type == "loess") {
    mod <- loess(Y ~ X, data = p2_df, weights = w2,
                 control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    mod <- randomForest(Y ~ X, data = p2_df)
  } else if (type == "spline") {
    mod <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2_df, weights = w2)
  } else if (type == "linear") {
    mod <- lm(Y ~ X, data = p2_df, weights = w2)
  }

@@

  p1_df <- p1_df %>%
    left_join(p2_df, by = join_by(X, Y, pi1, pi2, w1, w2, del1, del2)) %>%
    mutate(del2a = ifelse(is.na(del2a), 0, del2a)) %>%
    mutate(del2b = ifelse(is.na(del2b), 0, del2b))

  # 1. Estimate mhat in both splits
  if (type == "loess") {
    moda <- loess(Y ~ X, data = p2a_df, weights = w2,
                  control = loess.control(surface="direct"))
    modb <- loess(Y ~ X, data = p2b_df, weights = w2,
                  control = loess.control(surface="direct"))
  } else if (type == "randomforest") {
    moda <- randomForest(Y ~ X, data = p2a_df)
    modb <- randomForest(Y ~ X, data = p2b_df)
  } else if (type == "spline") {
    moda <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2a_df, weights = w2)
    modb <- gam(Y ~ s(X, k = 15, bs = "cr"), data = p2b_df, weights = w2)
  } else if (type == "linear") {
    moda <- lm(Y ~ X, data = p2a_df, weights = w2)
    modb <- lm(Y ~ X, data = p2b_df, weights = w2)
  }

@@

@@

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "randomForest", "mgcv"),
          .options.RNG  = 2) %dorng% {

@@

library(ggplot2)

ggplot(pop_df, aes(x = X, y = Y)) + 
geom_point() +
geom_smooth(method = "lm", formula = y ~ x)

ggplot(p2_df, aes(x = X, y = Y)) + 
geom_point() +
geom_smooth(method = "lm", formula = y ~ x)
